Supporting Information for "Imperfect slope measurements drive overestimation in geometric cone model of lake and reservoir depth"

### 2021-12-07 manuscript/si_template_2019.tex
reorder appendix

% \affiliation{1}{First Affiliation}
% \affiliation{2}{Second Affiliation}
% \affiliation{3}{Third Affiliation}
% \affiliation{4}{Fourth Affiliation}

% \affiliation{=number=}{=Affiliation Address=}
%(repeat as many times as is necessary)

@@

            \item \textbf{Figure S2:} Hypsography classification by state

            \item \textbf{Figure S3:} Lake geometry comparison among shape and reservoir classes

            \item \textbf{Figure S4:} Lake geometry comparison among shape classes by depth and area

            \item \textbf{Figure S5:} Effect of calculation technique on in-lake/nearshore slope comparison.

            \item \textbf{Figure S6:} Spatial distribution of depth model residuals.

            \item \textbf{Figure S7:} Characteristics of lakes with bathymetry versus non-bathymetry lakes

            \item \textbf{Figure S8:} Lake characteristics comparison among shape and reservoir classes

            \item \textbf{Figure S9:} Proxy-proxy model fit showing predicted depth versus measured depth

            \item \textbf{Figure S10:} Comparison between reported depth and bathymetry depth

@@

      % \noindent\textbf{Introduction}
      %Type or paste your text here. The introduction gives a brief overview of the supporting information. You should include information %about as many of the following as possible (when appropriate):
      % 1. a general overview of the kind of data files;
      % 2. information about when and how the data were collected or created;
      % 3. a general description of processing steps used;
      % 4. any known imperfections or anomalies in the data.

      %\clearpage

      %Delete all unused file types below. Copy/paste for multiples of each file type as needed.
      %\noindent\textbf{Text S1.}
      %Type or paste text here. This should be additional explanatory text, such as: extended descriptions of results, full details of models, extended lists of acknowledgements etc.  It should not be additional discussion, analysis, interpretation or critique. It should not be an additional scientific experiment or paper.
      %
      %Repeat for any additional Supporting Text

      %%Enter Data Set, Movie, and Audio captions here
      %%EXAMPLE CAPTIONS

      % \noindent\textbf{Data Set S1.} %Type or paste caption here.
      %upload your dataset(s) to AGU's journal submission site and select "Supporting Information (SI)" as the file type. Following naming %convention: ds01.

      %Repeat for any additional Supporting data sets

      % \noindent\textbf{Movie S1.} %Type or paste caption here.
      %upload your movie(s) to AGU's journal submission site and select, "Supporting Information %(SI)" as the file type. Following naming convention: ms01.

      %Repeat any additional Supporting movies

      % \noindent\textbf{Audio S1.} %Type or paste caption here.
      %upload your audio file(s) to AGU's journal submission site and select "Supporting Information %(SI)" as the file type. Following naming convention: auds01.

      %Repeat for any additional Supporting audio files

      %%% End of body of article:
      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      %
      % Optional Notation section goes here
      %
      % Notation -- End each entry with a period.
      % \begin{notation}
      % Term & definition.\\
      % Second term & second definition.\\
      % \end{notation}
      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

      %% ------------------------------------------------------------------------ %%
      %%  REFERENCE LIST AND TEXT CITATIONS

      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      %
      %
      % \bibliography{lagosdepth}
      %
      % no need to specify bibliographystyle
      %
      % Note that ALL references in this supporting information file must also be referenced in the primary manuscript
      %
      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      % if you get an error about newblock being undefined, uncomment this line:
      %\newcommand{\newblock}{}

      % \bibliography{ uncomment this line and enter the name of your bibtex file here }

      %Reference citation instructions and examples:
      %
      % Please use ONLY \cite and \citeA for reference citations.
      % \cite for parenthetical references
      % ...as shown in recent studies (Simpson et al., 2019)
      % \citeA for in-text citations
      % ...Simpson et al (2019) have shown...
      % DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
      %
      %
      %...as shown by \citeA{jskilby}.
      %...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
      %...has been shown \cite<e.g.,>{jskilbye}.
      %...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
      %...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
      %
      % apacite uses < > for prenotes, not [ ]
      % DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
      %

      %% ------------------------------------------------------------------------ %%
      %
      %  END ARTICLE
      %
      %% ------------------------------------------------------------------------ %%
\end{article}
\clearpage

@@

\begin{figure}
      \noindent\includegraphics[width=0.85\textwidth]{../figures/01_hypsography-1}
      \caption{Hypsography classification by state. Numbers on panel labels indicate the percentage of lakes in each state with a convex versus a concave cross-section shape.}\label{figS2}
\end{figure}

\begin{figure}
      \noindent\includegraphics[width=0.85\textwidth]{../figures/01_geometry_grid-1}
      \caption{Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope. A dashed 1:1 line is shown for comparison. Cross-section shape and reservoir class plots are not identical because not all lakes had a reservoir classification exceeding a 0.75 probability confidence level.}\label{figS3}
\end{figure}

\begin{figure}
      \begin{center}
            \includegraphics[width=0.79\textwidth,keepaspectratio]{../figures/slope_by_area_depth-1}
      \end{center}
      \caption{Comparison between in-lake and nearshore slopes in concave and convex lakes of the same size and max depth. Categories are quantile bins ($<$ 25\%, 25-50\%, 50-75\%, and 75-100\%).}\label{figS4}
\end{figure}
\clearpage

\begin{figure}
      \begin{center}
            \includegraphics[width=\textwidth,keepaspectratio]{../figures/01_slope_compare-1}
      \end{center}
      \caption{Comparison between in-lake and nearshore slope using different calculation techniques. The techniques used in the main text analyses are bolded and the combination of these techniques (top-left corner) produces the strongest relationship between the two metrics. \texttt{slope\_mean} is the mean slope of all inlake or nearshore buffer points. \texttt{slope\_pnts} is the average slope (i.e. \texttt{slope\_pnt}) of all points at maximum depth. \texttt{slope\_online\_mean} is the mean pixel-to-pixel slope of each pixel lying on a straight line either from the single deepest point to the lake shoreline (in the case of inlake slope) or from the lake shoreline point extending to the buffer exterior (in the case of nearshore slope). \texttt{slopes\_online\_mean} is the same as \texttt{slope\_online\_mean} except it uses all inlake points at maximum depth.}\label{figS5}
\end{figure}
\clearpage

\begin{figure}
      \begin{center}
            \includegraphics[width=\textwidth,keepaspectratio]{../figures/00_map_resid-1}
      \end{center}
      \caption{Spatial distribution of depth model residuals.}\label{figS6}
\end{figure}
\clearpage

% \begin{figure}
%       \noindent\includegraphics[width=0.85\textwidth]{../figures/02_depth_model_importance-1}
%       \caption{Importance plot for random forest variables showing increase in mean square error. Higher values indicate greater importance to model predictions. See Equation 1 for a definition of geometric max depth. HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.}\label{figS7}
% \end{figure}

\begin{figure}
      \begin{center}\includegraphics[width=0.45\textwidth]{../figures/gg_effort-1}\end{center}
      \caption{Comparison between characteristics of lakes with bathymetry data against lakes with depth from other sources in the LAGOSUS-Depth product. The distance to urban area metric is calculated using data from the 2018 US Census Urban and Rural Classification.}\label{figS7}
\end{figure}

\clearpage
\begin{figure}
      \begin{center}\includegraphics[width=0.85\textwidth]{../figures/01_contrasts-1}\end{center}
      \caption{Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status.}\label{figS8}
\end{figure}

\clearpage
\begin{figure}
      \begin{center}\includegraphics[width=0.85\textwidth]{../figures/02_depth_model_fitted_allproxy-1}\end{center}
      \caption{Proxy-proxy model fit showing predicted depth versus measured depth.}\label{figS9}
\end{figure}

@@

\begin{figure}
      \begin{center}\includegraphics[width=0.73\textwidth]{../figures/lgnemanual-vs-bathy-depth-1}\end{center}
      \caption{Comparison between reported depth and depth extracted from bathymetry surfaces by US State where reported depths come from the LAGOSUS-Depth product. For this figure, no reported depth values originated from the same source as its corresponding bathymetry-derived value.}\label{figS10}
\end{figure}

@@

% % source at tables/01_predictors.Rmd
% \begin{table}
%       \caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States). Predictor variables for computing random forest offsets (Equation 2) are printed in bold face. Dashes (-) indicate an identical sample size among this study and that of the contiguous United States from the National Hydrography Dataset. The total number lakes is reported as n.} \label{tableS1}
%       \centering
%       \setlength\tabcolsep{1.5pt} % default value: 6pt
%       \begin{tabular}{lllll}
%             \hline
%             Variable                            & Median      & Q25         & Q75          & n            \\
%             \hline
%             Max depth (m)                       & 8.2 (7)     & 4.6 (3.7)   & 14 (12)      & 4820 (17700) \\
%             \textbf{Area (ha)}                  & 55 (33)     & 21 (11)     & 140 (100)    & 4820 (17700) \\
%             \textbf{Island area (ha)}           & 0 (0)       & 0 (0)       & 0.18 (0.076) & 4820 (17700) \\
%             \textbf{Perimeter (m)}              & 4400 (3500) & 2500 (1800) & 8100 (7300)  & 4820 (17700) \\
%             \textbf{Shoreline development}      & 1.7 (1.7)   & 1.4 (1.4)   & 2.1 (2.2)    & 4820 (17700) \\
%             \textbf{Elevation (m)}              & 300 (340)   & 180 (210)   & 400 (460)    & 4820 (17700) \\
%             \textbf{Watershed-lake ratio}       & 7.8 (10)    & 3.8 (4.4)   & 17 (29)      & 4820 (17700) \\
%             \textbf{Deepest point distance (m)} & 180 (-)     & 110 (-)     & 290 (-)      & 4820 (-)     \\
%             Mean deepest point distance (m)     & 140 (-)     & 87 (-)      & 230 (-)      & 4820 (-)     \\
%             \textbf{Visual center distance (m)} & 240 (-)     & 160 (-)     & 390 (-)      & 4820 (-)     \\
%             Inlake slope (m/m)                  & 0.05 (-)    & 0.02 (-)    & 0.08 (-)     & 4820 (-)     \\
%             Inlake slope online (m/m)           & 0.06 (-)    & 0.03 (-)    & 0.14 (-)     & 4800 (-)     \\
%             Inlake slopes (m/m)                 & 0.06 (-)    & 0.03 (-)    & 0.1 (-)      & 4820 (-)     \\
%             Inlake slopes online (m/m)          & 0.07 (-)    & 0.03 (-)    & 0.15 (-)     & 4800 (-)     \\
%             Mean inlake slope (m/m)             & 0.04 (-)    & 0.02 (-)    & 0.09 (-)     & 4820 (-)     \\
%             Nearshore mean slope (m/m)          & 0.08 (-)    & 0.05 (-)    & 0.11 (-)     & 4820 (-)     \\
%             Nearshore slope online (m/m)        & 0.08 (-)    & 0.04 (-)    & 0.13 (-)     & 4590 (-)     \\
%             Nearshore slopes online (m/m)       & 0.08 (-)    & 0.04 (-)    & 0.13 (-)     & 4540 (-)     \\
%             \hline
%       \end{tabular}
% \end{table}

>>>>

@@

            \item \textbf{Figure S2:} Effect of calculation technique on in-lake/nearshore slope comparison.

            \item \textbf{Figure S3:} Hypsography classification by state

            \item \textbf{Figure S4:} Lake geometry comparison among shape and reservoir classes

            \item \textbf{Figure S5:} Lake characteristics comparison among shape and reservoir classes

            \item \textbf{Figure S6:} Proxy-proxy model fit showing predicted depth versus measured depth

            \item \textbf{Figure S7:} Spatial distribution of depth model residuals.

            \item \textbf{Figure S8:} Characteristics of lakes with bathymetry versus non-bathymetry lakes

            \item \textbf{Figure S9:} Comparison between reported depth and bathymetry depth

            \item \textbf{Figure S10:} Lake geometry comparison among shape classes by depth and area

@@

\end{article}
\clearpage

@@

\begin{figure}
      \begin{center}
            \includegraphics[width=\textwidth,keepaspectratio]{../figures/01_slope_compare-1}
      \end{center}
      \caption{Comparison between in-lake and nearshore slope using different calculation techniques. The techniques used in the main text analyses are bolded and the combination of these techniques (top-left corner) produces the strongest relationship between the two metrics. \texttt{slope\_mean} is the mean slope of all inlake or nearshore buffer points. \texttt{slope\_pnts} is the average slope (i.e. \texttt{slope\_pnt}) of all points at maximum depth. \texttt{slope\_online\_mean} is the mean pixel-to-pixel slope of each pixel lying on a straight line either from the single deepest point to the lake shoreline (in the case of inlake slope) or from the lake shoreline point extending to the buffer exterior (in the case of nearshore slope). \texttt{slopes\_online\_mean} is the same as \texttt{slope\_online\_mean} except it uses all inlake points at maximum depth.}\label{figS2}
\end{figure}

@@

\begin{figure}
      \noindent\includegraphics[width=0.85\textwidth]{../figures/01_hypsography-1}
      \caption{Hypsography classification by state. Numbers on panel labels indicate the percentage of lakes in each state with a convex versus a concave cross-section shape.}\label{figS3}
\end{figure}

\begin{figure}
      \noindent\includegraphics[width=0.85\textwidth]{../figures/01_geometry_grid-1}
      \caption{Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope. A dashed 1:1 line is shown for comparison. Cross-section shape and reservoir class plots are not identical because not all lakes had a reservoir classification exceeding a 0.75 probability confidence level.}\label{figS4}
\end{figure}

\clearpage
\begin{figure}
      \begin{center}\includegraphics[width=0.85\textwidth]{../figures/01_contrasts-1}\end{center}
      \caption{Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status.}\label{figS5}
\end{figure}

@@

\begin{figure}
      \begin{center}\includegraphics[width=0.85\textwidth]{../figures/02_depth_model_fitted_allproxy-1}\end{center}
      \caption{Proxy-proxy model fit showing predicted depth versus measured depth.}\label{figS6}
\end{figure}

\begin{figure}
      \begin{center}
            \includegraphics[width=\textwidth,keepaspectratio]{../figures/00_map_resid-1}
      \end{center}
      \caption{Spatial distribution of depth model residuals.}\label{figS7}
\end{figure}
\clearpage

\begin{figure}
      \begin{center}\includegraphics[width=0.45\textwidth]{../figures/gg_effort-1}\end{center}
      \caption{Comparison between characteristics of lakes with bathymetry data against lakes with depth from other sources in the LAGOSUS-Depth product. The distance to urban area metric is calculated using data from the 2018 US Census Urban and Rural Classification.}\label{figS8}
\end{figure}

\clearpage
\begin{figure}
      \begin{center}\includegraphics[width=0.73\textwidth]{../figures/lgnemanual-vs-bathy-depth-1}\end{center}
      \caption{Comparison between reported depth and depth extracted from bathymetry surfaces by US State where reported depths come from the LAGOSUS-Depth product. For this figure, no reported depth values originated from the same source as its corresponding bathymetry-derived value.}\label{figS9}
\end{figure}

\begin{figure}
      \begin{center}
            \includegraphics[width=0.79\textwidth,keepaspectratio]{../figures/slope_by_area_depth-1}
      \end{center}
      \caption{Comparison between in-lake and nearshore slopes in concave and convex lakes of the same size and max depth. Categories are quantile bins ($<$ 25\%, 25-50\%, 50-75\%, and 75-100\%).}\label{figS10}
\end{figure}
\clearpage
